---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(reshape2)
library(tidyverse)
library(dagitty)
library(ggdag)
library(survival)
library(survminer)
library(GGally)
library(ggpubr)
library(lavaan)
```

```{r}
dat <- read.csv("S1Data.csv")
dat
```

```{r}
col_names <- names(dat)
for (i in 1:ncol(dat)) {
  hist(dat[,i], main = col_names[i], xlab = col_names[i])
}
```

Can't stand the "Pletelets" misspelling. Also convert all binary variables to categorical or boolean.

```{r}
col_names[col_names == "Pletelets"] <- "Platelets"
names(dat) <- col_names

dat <- dat %>%
  mutate(
    Gender = factor(Gender, labels = c("Female", "Male")),
    across(c(Event, Smoking:Anaemia), as.logical)
  )

dat
```


```{r fig.height=10, message=FALSE}
ggpairs(dat, aes(alpha = 0.01), progress = FALSE)
```

From the plot above, there seems to be a reasonably strong correlation between smoking and gender.

```{r}
ggplot(dat, aes(Gender, Smoking)) +
  geom_jitter()
```

```{r}
chisq.test(dat$Gender, dat$Smoking)
```


# Find cut-off point

Apparently barely any women in our data set were smoking.

```{r}
fit <- survfit(Surv(TIME, Event) ~ Gender, data = dat)
fit
```

```{r fig.height=8}
ggsurvplot(fit,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"),
          ncensor.plot = TRUE)
```

```{r}
cut_offs <- 0:200
names(cut_offs) <- paste0("cut_off", cut_offs)

dat_status <- dat %>%
  add_column(!!!cut_offs) %>%
  pivot_longer(cols = starts_with("cut_off"), 
               names_to = "cut_off", 
               values_to = "cut_off_val") %>%
  group_by(cut_off_val) %>%
  mutate(
    censored = TIME < cut_off_val & Event == 0,
    deceased = TIME < cut_off_val & Event == 1,
    confirmed_alive = TIME >= cut_off_val
  ) %>%
  summarize(
    n_censored = sum(censored),
    n_deceased = sum(deceased),
    n_confirmed_alive = sum(confirmed_alive)
  ) 

dat_status
```

```{r}
df <- pivot_longer(dat_status, n_censored:n_confirmed_alive, names_to = "status")

ggplot(df, aes(cut_off_val, value)) +
  geom_line(aes(color = status), linewidth = 1) +
  labs(x = "Time (day)", y = "Number of patients") +
  scale_color_hue(name = "Patient status", labels = c("Censored", "Alive", "Deceased")) +
  theme_minimal()

ggsave(file.path("images", "cut_off.pdf"))
```

There are periods of a sharp increase in censored patients, while the number of deceased patients only climbs gradually. The first jump in censored patients is just after t = 75, so around that time might be a good cut-off point. From the dat_status table, it is clear that we should then pick t = 74, as the only difference between t = 74 and t = 75 is an increase in censored patients.

```{r}
dat_cut <- dat %>%
  filter(!(TIME < 74 & Event == 0)) %>%
  rename(Deceased = Event) %>%
  select(!TIME)

dat_cut
```

```{r}
saveRDS(dat_cut, "dat_cut.rds")
```

For convenience:

```{r}
dat_cut <- readRDS("dat_cut.rds")
```

# Prepare Bayesian network

```{r paged.print=FALSE}
bn <- dagitty("dag{
    Gender -> {Smoking Creatinine Ejection.Fraction}
    Age -> {Creatinine Deceased Ejection.Fraction BP}
    BP -> {Creatinine Deceased Ejection.Fraction}
    Sodium -> Ejection.Fraction
    Ejection.Fraction -> Deceased
    Creatinine -> Deceased
    Smoking -> {Creatinine Deceased Ejection.Fraction BP}
  }")

# Just for plotting
bn_plt <- tidy_dagitty(bn, layout = "circle")
node_labels <- c(
  "Age" = "Age", 
  "BP" = "BP", 
  "Creatinine" = "SC", 
  "Deceased" = "D", 
  "Ejection.Fraction" = "EF", 
  "Gender" = "Sex", 
  "Smoking" = "Sm", 
  "Sodium" = "SS"
)
bn_plt <- bn_plt %>%
  mutate(
    name = node_labels[name],
    to = node_labels[to]
  )

ggdag(bn_plt) +
  theme_dag_blank()

ggsave(file.path('images', 'network_literature.pdf'))
```

# Test network

```{r}
impliedConditionalIndependencies(bn)
```

## Correlations

Convert all boolean/categorical columns to numeric, then run `localTests` with `type = "cis"`. This will regress all variables on their parents, and use the correlation between the resulting residuals to assess the amount of remaining dependence between the regressed variables.

```{r paged.print=FALSE}
dat_cut_num <- dat_cut %>%
  mutate(across(where(~ !is.numeric(.x)), as.numeric))

localTests(bn, dat_cut_num, type = "cis")
```

Creatinine and sodium have a relatively high correlation estimate and low p value, so they might not be independent after all. It is not the only violated independence assumption, but let's start by adding an edge from creatinine to sodium.

```{r}
bn2 <- dagitty("dag{
    Gender -> {Smoking Creatinine Ejection.Fraction}
    Age -> {Creatinine Deceased Ejection.Fraction BP}
    BP -> {Creatinine Deceased Ejection.Fraction}
    Sodium -> Ejection.Fraction
    Ejection.Fraction -> Deceased
    Creatinine -> {Deceased Sodium}
    Smoking -> {Creatinine Deceased Ejection.Fraction BP}
  }")

ggdag(bn2, layout = "circle")
```

And run our tests again:

```{r paged.print=FALSE}
localTests(bn2, dat_cut_num, type = "cis")
```

This resolved all other independence assumption violations.

## Canonical correlations

Estimates are exactly the same as in the vanilla correlation case, p values are slightly different.

```{r paged.print=FALSE}
localTests(bn, dat_cut, type = "cis.pillai")
```

```{r}
bn2 <- dagitty("dag{
    Gender -> {Smoking Creatinine Ejection.Fraction}
    Age -> {Creatinine Deceased Ejection.Fraction BP}
    BP -> {Creatinine Deceased Ejection.Fraction}
    Sodium -> Ejection.Fraction
    Ejection.Fraction -> Deceased
    Creatinine -> {Deceased Sodium}
    Smoking -> {Creatinine Deceased Ejection.Fraction BP}
  }")

ggdag(bn2, layout = "circle")
```

```{r paged.print=FALSE}
localTests(bn2, dat_cut, type = "cis.pillai")
```

# Find network coefficients

## Canonical correlations

```{r}
r <- c()
for (n in names(bn2)) {
  for (p in parents(bn2, n)) {
    otherparents <- setdiff(parents(bn2, n), p)
    tst <- ciTest(X=n, Y=p, Z=otherparents, dat_cut, type="cis.pillai" )
    r <- rbind(r, data.frame(
      name=p,
      direction="->",
      to=n, 
      cor=tst[,"estimate"],
      p=tst[,"p.value"]
    ))
  }
}

r
```

```{r}
bn2_tidy <- tidy_dagitty(bn2, layout = "circle") 

bn2_tidy$data <- bn2_tidy$data %>%
  full_join(r, by = c("name", "direction", "to")) %>%
  mutate(
    x_text = (x + xend) / 2, 
    y_text = (y + yend) / 2,
    cor = round(cor, 2),
    # Some manual adjustments to certain labels clearer
    x_text = case_when(
      name == "Age" & to == "Ejection.Fraction" ~ 0.1,
      name == "BP" & to == "Deceased" ~ -0.354,
      TRUE ~ x_text
    ),
    y_text = case_when(
      name == "Smoking" & to == "Creatinine" ~ -0.2,
      TRUE ~ y_text
    )
  )

bn2_tidy
```

```{r}
ggdag(bn2_tidy) +
  geom_label(aes(x_text, y_text, label = cor), 
             data = filter(bn2_tidy, abs(cor) > 0.01))
```

## Polychoric correlations

Here we need to treat binary variables as numeric. Perhaps it makes more sense to estimate the GGM using linear regressions rather than polychoric correlations?

```{r}
dat_cut_num <- dat_cut %>%
  mutate(
    across(c(Deceased:Anaemia), as.numeric),
    across(everything(), scale) # Otherwise lavaan complains
  )

cov_mat <- lavCor(dat_cut_num)
cov_mat
```

```{r}
localTests(bn2, sample.cov = cov_mat, sample.nobs = nrow(dat_cut_num))
```

```{r}
cov_df <- melt(cov_mat) %>%
  rename(
    name = Var1,
    to = Var2,
    poly_cor = value
  )

bn2_tidy_poly <- bn2_tidy
bn2_tidy_poly$data <- bn2_tidy_poly$data %>%
  left_join(cov_df, by = c("name", "to")) %>%
  mutate(poly_cor = round(poly_cor, 2))

bn2_tidy_poly
```

```{r}
ggdag(bn2_tidy_poly) +
  geom_label(aes(x_text, y_text, label = poly_cor), 
             data = filter(bn2_tidy_poly, abs(poly_cor) > 0.01))
```

## Gaussian graphical model

```{r}
ggm <- "
  Creatinine ~ Gender + Smoking + Age + BP
  BP ~ Smoking + Age
  Sodium ~ Creatinine
  Smoking ~ Gender
  Ejection.Fraction ~ BP + Age + Sodium + Smoking + Gender
  Deceased ~ Creatinine + BP + Age + Smoking + Ejection.Fraction
"

ggm_sem <- sem(ggm, data = dat_cut_num)
ggm_sem
```

```{r}
summ <- summary(ggm_sem)
summ
```

```{r}
ggm_est <- summ$pe %>%
  filter(op == "~") %>%
  select(lhs, rhs, est, pvalue) %>%
  rename(sem_est = est, sem_pvalue = pvalue) %>%
  mutate(sem_est = round(sem_est, 2))

ggm_est
```

```{r}
bn2_tidy_sem <- bn2_tidy

bn2_tidy_sem$data <- bn2_tidy_sem$data %>%
  left_join(ggm_est, by = c("name" = "rhs", "to" = "lhs"))

bn2_tidy_sem
```

```{r}
ggdag(bn2_tidy_sem) +
  geom_label(aes(x_text, y_text, label = sem_est), 
             data = filter(bn2_tidy_sem, abs(sem_est) > 0.01))
```

Remove edges with p-values below 0.05.

```{r}
bn3_tidy_sem <- bn2_tidy_sem
bn3_tidy_sem$data <- bn3_tidy_sem$data %>%
  filter(sem_pvalue < 0.05) %>%
  # Add deleted nodes back in
  add_row(
    name = c("Deceased", "Smoking"),
    x = c(-7.071068e-01, 0),
    y = c(7.071068e-01, -1)
  )

ggdag(bn3_tidy_sem) +
  geom_label(aes(x_text, y_text, label = sem_est))
```

### Trimmed GGM

Rerun SEM with trimmed graphical model.

```{r}
ggm2 <- "
  Creatinine ~ Age
  Sodium ~ Creatinine
  Smoking ~ Gender
  Ejection.Fraction ~ Sodium + Gender
  Deceased ~ Creatinine + Age + Ejection.Fraction
  Smoking ~~ 0*Deceased
  Age ~~ 0*Gender
"

ggm_sem2 <- sem(ggm2, data = dat_cut_num)
ggm_sem2
```

```{r}
summary(ggm_sem2)
```

```{r}
summary(ggm_sem2)$pe
```


```{r}
ggm2_coef <- summary(ggm_sem2)$pe

ggm2_var <- ggm2_coef %>%
  filter(lhs == rhs) %>%
  select(rhs, est) %>%
  rename(var = est) %>%
  mutate(var = round(var, 2))

ggm_est2 <- ggm2_coef %>%
  filter(op == "~") %>%
  select(lhs, rhs, est, pvalue) %>%
  rename(sem_est2 = est, sem_pvalue2 = pvalue) %>%
  mutate(sem_est2 = round(sem_est2, 2)) %>%
  left_join(ggm2_var, by = c("rhs"))
  
ggm_est2
```

Taking just a quick look at the new coefficients. Comparing them with the first GGM shows that they have not changed. Too lazy to fix the missing Smoking node.

```{r}
bn4_tidy_sem <- bn2_tidy_sem
bn4_tidy_sem$data <- bn4_tidy_sem$data %>%
  left_join(ggm_est2, by = c("name" = "rhs", "to" = "lhs"))

ggdag(bn4_tidy_sem %>% filter(is.na(to) | !is.na(sem_est2))) +
  geom_label(aes(x_text, y_text, label = sem_est2))
```

## Cox regression

We are keeping in all edges as determined from the literature, even if the GGM would have them removed.

```{r}
# Scale dat to make sure coefficients have same scale
dat_scaled <- dat %>%
  mutate(across(where(is.numeric), function(x) {
    if (cur_column() == "TIME") return(x)
    scale(x)
  }))

cox_m1 <- coxph(Surv(TIME, Event) ~ Creatinine + 
                  Age + 
                  Smoking + 
                  Ejection.Fraction +
                  BP,
                data = dat_scaled)
cox_m1
```

```{r}
summ_cox <- summary(cox_m1)
summ_cox
```

This shows that smoking barely has any effect on survival. Removing the edge between Smoking and Deceased would cause Smoking to no longer have any outgoing edges, so we drop it completely from the network.

```{r}
cox_m1_fit <- survfit(cox_m1, data = dat_scaled)

# Plot the baseline survival function
ggsurvplot(cox_m1_fit, palette = "#2E9FDF",
           ggtheme = theme_minimal())
```

```{r}
############# Serum creatinine

crea_df <- with(dat_scaled, {
  data.frame(
    Age = mean(Age),
    Smoking = FALSE,
    Ejection.Fraction = mean(Ejection.Fraction),
    BP = FALSE,
    Creatinine = quantile(Creatinine, c(0.1, 0.5, 0.9))
  )
})

g_crea <- ggsurvplot(survfit(cox_m1, newdata = crea_df),
           ggtheme = theme_minimal(), data = crea_df) +
  labs(title = "A) Survival by serum creatinine levels")

############# Ejection fraction

ejec_df <- with(dat_scaled, {
  data.frame(
    Age = mean(Age),
    Smoking = FALSE,
    Ejection.Fraction = quantile(Ejection.Fraction, c(0.1, 0.5, 0.9)),
    BP = FALSE,
    Creatinine = mean(Creatinine)
  )
})

g_ejec <- ggsurvplot(survfit(cox_m1, newdata = ejec_df),
           ggtheme = theme_minimal(), data = ejec_df) +
  labs(title = "B) Survival by ejection fraction levels")

############# Age

# Roughly corresponds to 10%, 50%, 90% quantiles
age_scale <- attr(dat_scaled$Age, "scaled:scale")
age_center <- attr(dat_scaled$Age, "scaled:center")
ages = c(45, 60, 75)

age_df <- with(dat_scaled, {
  data.frame(
    Age = (ages - age_center) / age_scale,
    Smoking = FALSE,
    Ejection.Fraction = mean(Ejection.Fraction),
    BP = FALSE,
    Creatinine = mean(Creatinine)
  )
})

cox_m1_fit_age <- survfit(cox_m1, newdata = age_df)
g_age <- ggsurvplot(cox_m1_fit_age, legend.labs = ages,
                    ggtheme = theme_minimal(), data = age_df) +
  labs(title = "C) Survival by age")

############# BP

bp_df <- with(dat_scaled, {
  data.frame(
    Age = mean(Age),
    Smoking = FALSE,
    Ejection.Fraction = mean(Ejection.Fraction),
    BP = c(TRUE, FALSE),
    Creatinine = mean(Creatinine)
  )
})

g_bp <- ggsurvplot(survfit(cox_m1, newdata = bp_df, legend.labs = c("TRUE", "FALSE")),
                   ggtheme = theme_minimal(), data = bp_df)  +
  labs(title = "D) Survival by high blood pressure")

g_bp$plot <- g_bp$plot +
  scale_color_hue(labels = c("High", "Low")) +
  scale_fill_hue(labels = c("High", "Low"))
```

```{r}
quantile(dat$Creatinine, c(0.1, 0.5, 0.9))
quantile(dat$Ejection.Fraction, c(0.1, 0.5, 0.9))
```


```{r fig.height=10}
ggarrange(plotlist = lapply(list(g_crea, g_ejec, g_age, g_bp), function (sp) {
  sp$plot +
    theme(
      text = element_text(size = 12),
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    )
}))
ggsave(file.path("images", "surv_strata.pdf"), width = 12, height = 10)
```


## Final network

Drop Smoking and merge Cox regression estimates with second set of GGM estimates.

```{r}
summ_cox$coefficients[,1]
```


```{r}
cox_coef <- summ_cox$coefficients[,1]
names(cox_coef) <- c("Creatinine", "Age", "Smoking", "Ejection.Fraction", "BP")

# Recreate tidy DAG to get a new layout
bn_final <- tidy_dagitty(bn2, layout = 'dh')

bn_final$data <- bn_final$data %>%
  left_join(ggm_est2, by = c("name" = "rhs", "to" = "lhs")) %>%
  mutate(
    coef = case_when(
      to == "Deceased" & name %in% names(cox_coef) ~ round(cox_coef[name], 2),
      TRUE ~ sem_est2
    ),
    var = case_when(
      var != 1 ~ var,
      TRUE ~ NA_real_
    )
  ) %>%
  relocate(coef, .before = xend) %>%
  filter((name != "Smoking" & to != "Smoking" & !is.na(coef)) | is.na(to)) %>%
  mutate(
    name = node_labels[name],
    to = node_labels[to]
  )

bn_final

ggplot(bn_final, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point() +
  geom_dag_edges(aes(label = coef), 
                 angle_calc = "along", 
                 label_dodge = unit(2.5, 'mm')) +
  geom_dag_text(col = "white")
```

Perhaps it's clearer if we fix the coordinates ourselves:

```{r}
coords <- matrix(c(
  -10, -11, # Age
  -7, -5, # BP
  -10, -9, # SC
  -7, -8, # D
  -10, -5, # EF
  -13, -5, # Sex
  0, 0, # Smoking [not included]
  -10, -7 # So
), ncol = 2, byrow = TRUE)

rownames(coords) <- node_labels
colnames(coords) <- c("x", "y")

bn_tmp <- bn_final$data %>%
  mutate(
    x = coords[name, "x"],
    y = coords[name, "y"],
    xend = coords[replace_na(to, "Sm"), "x"],
    yend = coords[replace_na(to, "Sm"), "y"]
  )

ggplot(bn_tmp, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point() +
  geom_dag_edges(aes(label = coef), 
                 angle_calc = "along", 
                 label_dodge = unit(-2.75, 'mm')) +
  geom_dag_text(col = "white") +
  geom_label(aes(label = ifelse(is.na(var), NA, paste("sigma^2 ==", var))), 
             nudge_x = -0.4, nudge_y = 0.7, parse = TRUE, hjust = 1) +
  xlim(-15, -5) +
  ylim(-11.5, -3.5) +
  theme_dag_blank()

ggsave(file.path("images", "network_coefficients.pdf"))
```



